Unlocking Sensitivity with A Two-baseline Power Spectrum Estimator for 21\,cm Inteferometric Probes of Reionization

### 2017-07-14 Rotation paper/draft_working.tex
saving progress

We make the distinction between a \textit{baseline}, which corresponds to two specific antennas, and a \textit{class of baselines}, which refers to all baselines of the same length and orientation in a given array. 
Baselines of the same class are traditionally called
``redundant baselines'', because they measure the same Fourier mode
in the sky.  We shall call baselines in the same class \textit{equivalent baselines}, and reserve the word \textit{redundancy} of two baselines in reference to a variable function of the relative time-offsets of their visibility time series. With this terminology, two equivalent
baselines are fully redundant with each other simultaneously at all
times. Non-equivalent baselines also have partial redundancy, and the redundancy can be maximized if their
respected time series are shifted with respect to one another by some delay.

@@

We shall use the 128-element PAPER array (henceforth as PAPER-128) to motivate our formalism and demonstrate our method, and extend our results to several HERA configurations in Sec \ref{sec:arrconf}. 
The PAPER array is located in the Karoo desert in South Africa (30:43:17.5
S, 21:25:41.8 E). The layout pattern with antenna labels is shown
in Fig. \ref{fig:AntPos}. The array consists of a 112-element core is a rectangular grid, and 16 "outriggers" used primarily to aid calibration.  In the bottom panel, the two baselines marked in red, are an example of equivalent pair. We denote a equivalency-class of baselines in the PAPER grid by their separations, in this case  \{2,0\}, for the
antennas are separated by 1 unit east and 0 unit north. Similarly,
the baselines marked in blue and yellow are respectively examples
of \{2,1\} and \{2, -1\}.
Note \{2,0\} and \{-2, 0\} for example are the same class and should
not be counted twice. Antennas in purely north-south baselines
are close (4m), and hence these baselines are not suitable
for cross-multiplication due to cross-coupling. On the other hand, the close North-South separation means that classes such as \{2,0\} and  \{2,1\} are expected to be near-equivalent. The PAPER-64 analysis of \cite{Ali2015} used three classes of baselines, the PAPER-128
equivalent of which are 
\{2, 0\}, \{2, 1\} and \{2, -1\}. There each of these classes
of baselines are cross multiplied within itself. This paper provides the method for inter-class multiplications. To do so we will use the short hand notation $\{m,n\}:\{m',n'\}$ to denote a pair of baseline classes to be cross-multiplied.

@@

Radio interferometric observations are often described in the coordinates $uvw$ define as:
\[
(u, v, w) = \frac{\nu}{c}\boldsymbol{b}, 
\]
where $\boldsymbol{b}$ is the baseline vector in Cartesian coordinates, with first and second coordinates pointing East and North, respectively.  $\nu$ is the frequency of observation, and $c$ is the speed of light. 
Relative to a phase center on the sky, each baseline maps to a point in $uvw$ space. As the Earth rotates, the points trace out tracks in the $uvw$ space. 
We show in Fig. \ref{fig:Tracks} $uvw$ tracks of the 3 PAPER-128 baselines colored in Fig. \ref{fig:AntPos}, projected onto $uv$ and $vw$ planes. The tracks are traced over 12 sidereal hours, at 0.15GHz, relative to a phase center that passes through zenith.

\begin{figure}[h]
\includegraphics[width=\linewidth]{rotation_new}
\caption{Tracks of 3 PAPER-128 baseline classes shown in Fig. \ref{fig:AntPos}, here with the same respective colors. Tracks are traced out over half a sidereal day for frequency $\nu=0.15\text{GHz}$ and a phase center that passes through zenith. The top panel shows $uv$ tracks with crossings among each pair of baseline classes. Projection to the $vw$ plane in the bottom panel shows that tracks in $uvw$ space do not cross. }
\label{fig:Tracks}
\end{figure}

Equivalent baselines follow identical $uvw$ tracks. Traditionally, we can identify
redundancy of nearly equivalent baselines as crossings
of the $uv$ tracks, a 2 dimensional projection of Fig. \ref{fig:Tracks}. As we see in the top panel of Fig. \ref{fig:Tracks}, there would be many such crossings. However, there are several reasons that $uv$ track-crossing do not imply perfect redundancy. In fact, in our case $uv$ track-crossings are
not accurate enough for time offset determination, nor can it give estimate of the degree of redundancy.  The most obvious reason is that the 3 dimensional $uvw$ tracks do not actually cross, as is evident from the $vw$ projection in the bottom panel of Fig. \ref{fig:Tracks}. For drift-scan arrays, even a hypothetical crossing in $uvw$ space would not imply perfect redundancy. This shall become evident in the next section, after we develop a more general formalism that accounts for the point spread function of the finite beams, and estimate the degree of redundancy for general combination of baselines at a general time-offset. Then, relation of track-crossing and redundancy for both drift-scan and tracking measurements will be explored in Appendix \ref{sec:appA}.

@@

%\caption{Example beam response of HERA (left) and PAPER (right) antennas. Both beams are Stokes $I$ polarization antenna voltage beams, at frequency of $\nu=150\text{MHz}$ and normalized to $1$ at zenith. The circles centered around zenith (center of beam) here are
%spaced 10 degrees apart. \label{fig:Beam}}
%\end{figure}

Power-spectrum measurements are typically taken from $\sim10$MHz centered around the corresponding redshift of interest (e.g. 150 MHz for z=9.5).

@@

Ultimately,
we would like to relate the data, collected with coordinates $\hat{\boldsymbol{s}}$
and $\nu$, to the power spectrum, written with 
$\boldsymbol{r}$ and $\boldsymbol{k}$, the cosmological position coordinate and wavenumber. To do so, we start by noticing that
\[
\begin{aligned}r & =\frac{c}{H_{0}}\int_{0}^{z}\frac{dz'}{E(z')},\\
 & \approx\frac{c}{H_{0}}\int_{0}^{z_{0}}\frac{dz'}{E(z')}-\frac{c(1+z)^{2}}{\nu_{21}H_{0}E(z)}\left(\nu-\nu_{0}\right),\\
 & \equiv X-Y\Delta\nu,
\end{aligned} \label{eq:r}
\]
where $\nu_{21}=1420$MHz is the 21\,cm transition rest frequency, $\nu_{0}$
a reference central frequency with corresponding redshift $z_{0}$,
and 
\[
E(z)=\sqrt{\Omega_{m}(1+z)^{3}+\Omega_{\Lambda}}.
\]
Inverting for $\nu$:
\begin{equation}
\nu=\frac{X-r}{Y}+\nu_{21}.\label{eq:nur}
\end{equation}

@@

Existing visibility based power-spectrum pipelines for redundant drift-scan  arrays relate the power-spectrum to the conjugate square of the visibilities \citep{delay-transform, paper32, Ali2015}. We would like to generalize such relations by relating the power-spectrum to the product of two visibilities from two arbitrary baselines and time-offsets. 
The beam pattern of a baseline shifts relative to the sky as the Earth rotates. Here we choose to fix the sky, and denote the rotated coordinates
in the topocentric frame with the 3 dimensional rotation operator $\boldsymbol{\Gamma}$:

@@

With implicit bounds of integrals from $-\infty$ to $\infty$, we have:
\begin{widetext}
\begin{equation}
\begin{aligned} \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle
 & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),\\
 & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),\\
 & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),\\
 & \approx P(k_{b,\tau})\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2} \delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),\\
 & = P(k_{b,\tau})\int \frac{d^{3}r}{(X^2Y)^2}B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r})\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right],\\
 & = P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}B^{*}(\hat{\boldsymbol{s}},\nu)B(\boldsymbol{\Gamma} \hat{\boldsymbol{s}},\nu)\exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right],
\end{aligned}
\label{eq:main}
\end{equation}

@@

\begin{equation}
\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle\\
 & \qquad \approx P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}B^{*}(\hat{\boldsymbol{s}},\nu)B(\boldsymbol{\Gamma}\hat{\boldsymbol{s}},\nu)\\
 & \qquad \qquad \qquad \qquad \exp\left[i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right].
 \end{aligned}
\label{eq:final}
\end{equation}

@@

\begin{equation}
 \hat{P}(k_{b,\tau}) \equiv \frac{V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)}{\Theta}, 
 \label{eq:opp}
\end{equation}
where the weight is defined as 
\begin{equation}
\Theta \equiv\int d\nu \Theta_{\nu}, 
\label{eq:Theta}
\end{equation}
with

@@

To first order, the physical origin of $\psi_\nu$ lies in the two visibilities having different phase centers. By default the correlators of a drift-scan array phase the two visibilities both to zenith at the same time. When they are cross-multiplied with a time lag, the visibilities must be rephased before delay transform to the account for the movement of the zenith. The effect of the phase thus roughly corresponds in a shift of the delay mode measured, and we should expect $\psi_{\nu}$ to be in first order a linear function of $\nu$:
\[
\psi_{\nu}\sim 2\pi\Delta\tau\nu.
\]
In Fig. \ref{fig:phi_nu} we show the peak phases for for a given baseline pair of PAPER128 (\{2,0\}:\{2,1\}), comparing the drift-scan phase dependence with that of the same baseline with hypothetical tracking-elements (Eq. \eqref{eq:Thetanu} and Eq. \eqref{eq:Thetanu_tracking}). In the drift-scan case, we indeed see a linear relation corresponding to a delay of $\Delta\tau\approx15.7$ns. In the tracking case, because of minimal zenith movement, only second order effects are observed. The origin of the second order effects can be seen as due to the $w$-term, or more precisely the fact that $uvw$ tracks do not cross when their 2 dimensional projections do. We refer the reader to Appendix \ref{sec:appA} for further explanation. In both cases, the full effects are encapsulated in the phase of $\Theta_\nu$ and can be thus determined empirically without any additional computation.

@@

\caption{Frequency dependent peak phases for PAPER-128 baseline pair \{2,0\}:\{2,1\}. Both the drift-scan case and a hypothetical tracking baseline are shown in solid lines. The first-order, linear effects are shown in dashed lines. We have also fixed the global phase in both cases to 0 at 0.15GHz. The drift-scan case exhibits linear behavior and conforms well to the first order effect of a shift in delay space due to the movement of zenith. }
\label{fig:phi_nu}
\end{figure}

To illustrate the effect of rephasing, we compare in Fig. \ref{fig:freqdiff} real parts of $\Theta_{\nu}$ of both equivalent and nearly equivalent baseline pairs for two channels: 0.145 GHz and 0.155 GHz. The top two panels have zero rephasing, and the bottom two are linearly rephased to a differential delay of $\Delta\tau\approx15.7$ns, as we determined above. The first and third panels show the equivalent baseline pairs \{2,0\}:\{2,0\}, and second and fourth panels show \{2,0\}:\{2,1\}. Summing over frequency without proper rephasing leads to destructive interference and signal loss. The wider the frequency profile, the more destructive the interference would be. Only after rephasing to the correct time-offset for each baseline pairs can we constructively combine the frequency channels.

@@

So far for clarity and generality we have avoided the traditional formulation of radio astronomy in terms of the $uv$-plane. To gain some visual intuition of the formalism in Section \ref{sec:method}, we show explicit beam fringe patterns of the two baselines and their interactions. For this section, we relax into the flat sky approximation so that $(l,m,\sqrt{1-l^2-m^2})\equiv\hat{s}$ and examine a single frequency of $150$MHz.

At the time-offset of maximum redundancy, we expect two beams to have to same fringe pattern in both frequency and phase. Due to the time delay, however, the beam centers would be slightly shifted with respect to each other. This we show in the top panels of Fig. \ref{fig:beamfringe}. The top left and middle panels show the beam-fringe
patterns (real parts) for baselines \{2,0\} and \{2,1\}, delayed by 0.0325 sidereal days.
The product of the beam fringe patterns in top right shows that the fringes
 cancel out as we expect. On the bottom of Fig. \ref{fig:beamfringe} we show the instantaneous $uv$ coverage of two baselines and their product. We see from the bottom right panel that the cross-multiplied baselines recover  power concentrated at $u\approx\frac{30\text{m}}{c/1.5\times10^8\text{Hz}}\approx15$. Note the bottom right panel is not the Fourier transform of the top right panel.

@@

\caption{PAPER128 beam-fringe patterns and their point-spread functions on the $uv$-plane for a frequency of $150$MHz. Selected baselines have 30 meter East-West separation, and 0 and 4 meter North-South separation, respectively. The top panels show the real parts of the beam-fringe patterns of the baselines (left and middle), and their conjugate product (right). The beam-fringe values are normalized such that the peak of the original beam is unity. 
The bottom left and middle panels show the peak-normalized $uv$ beam power of the two beam-fringe patterns. The product of the beam spreads displayed in the bottom right panel shows that power concentrated at $u\approx15$ is recovered. }
\label{fig:beamfringe}
\end{figure*}

@@

Bringing together the discussion from Section \ref{sec:method}, we present a numerical check of Eq. \eqref{eq:final}, including rephasing. To do so, we need to compare the amplitude and phase of the integral weight $\Theta$ 
for a pair of baselines with products of simulated visibilities of those baselines. We use 10 frequency channels evenly spaced from $145$MHz to $155$MHz for the comparison. We rephase one of the simulated visibilities, as well as computed $\Theta_{\nu}$ to the same delay as computed from the peak phase of the latter.

@@

Fig. \ref{fig:numerics} shows the real parts of the resulting degree of redundancy as a function of time offset, for equivalent baseline pairs (\{2,0\}:\{2,0\}) in the top panel and nearly equivalent (\{2,0\}:\{2,1\}) in the bottom. Shown in green for the two cases are the computed values of $\Theta$, both normalized to the peak of the equivalent case. The plot on the bottom is rephased with a delay of $\Delta\tau\approx15.7$ns as outlined in Section \ref{sec:rephs}. Since both y-axes are normalized to the same scale, we see the equivalent baseline reach maximal redundancy with 0 time-offset, while the nearly equivalent pair reaches maximum redundancy of $\Theta\approx0.5$ at $dT\approx0.325$days.

>>>>

We make the distinction between a \textit{baseline}, which corresponds to two specific antennas, and a \textit{class of baselines}, which refers to all baselines of the same length and orientation in a given array. 
Baselines of the same class are traditionally called
``redundant baselines'', because they measure the same Fourier mode
in the sky.  We shall call baselines in the same class \textit{equivalent baselines}, and reserve the word \textit{redundancy} of two baselines in reference to a variable function of the relative time offsets of their visibility time series. With this terminology, two equivalent
baselines are fully redundant with each other simultaneously at all
times. Non-equivalent baselines also have partial redundancy, and the redundancy can be maximized if their
respected time series are shifted with respect to one another by some delay.

@@

We shall use the 128-element PAPER array (henceforth as PAPER-128) to motivate our formalism and demonstrate our method, and extend our results to several HERA configurations in Sec \ref{sec:arrconf}. 
The PAPER array is located in the Karoo desert in South Africa (30:43:17.5
S, 21:25:41.8 E). The layout pattern with antenna labels is shown
in Fig. \ref{fig:AntPos}. The array consists of a 112-element core is a rectangular grid, and 16 "outriggers" used primarily to aid calibration.  In the bottom panel, the two baselines marked in red, are an example of equivalent pair. We denote a equivalency-class of baselines in the PAPER grid by their separations, in this case  \{2,0\}, for the
antennas are separated by 1 unit east and 0 unit north. Similarly,
the baselines marked in blue and yellow are respectively examples
of \{2,1\} and \{2, -1\}.
Note \{2,0\} and \{-2, 0\} for example are the same class and should
not be counted twice. Antennas in purely north-south baselines
are close (4m), and hence these baselines are not suitable
for cross-multiplication due to cross-coupling. On the other hand, the close North-South separation means that classes such as \{2,0\} and  \{2,1\} are expected to be near-equivalent. The PAPER-64 analysis of \cite{Ali2015} used three classes of baselines, the PAPER-128
equivalent of which are 
\{2, 0\}, \{2, 1\} and \{2, -1\}. There each of these classes
of baselines are cross multiplied within itself. This paper provides the method for inter-class multiplications. To do so we will use the short hand notation $\{m,n:m',n'\}$ to denote a pair of baseline classes to be cross-multiplied.

@@

Radio interferometric observations are often described in the coordinates $uvw$ define as:
\[
(u, v, w) = \frac{\nu}{c}\boldsymbol{b}, 
\]
where $\boldsymbol{b}$ is the baseline vector in Cartesian coordinates, with first and second coordinates pointing East and North, respectively.  $\nu$ is the frequency of observation, and $c$ is the speed of light. 
Relative to a phase center on the sky, each baseline maps to a point in $uvw$ space. As the Earth rotates, the points trace out tracks in the $uvw$ space. 
We show in Fig. \ref{fig:Tracks} $uvw$ tracks of the 3 PAPER-128 baselines colored in Fig. \ref{fig:AntPos}, projected onto $uv$ and $vw$ planes. The tracks are traced over 12 sidereal hours, at 0.15\,GHz, relative to a phase center that passes through zenith.

\begin{figure}[h]
\includegraphics[width=\linewidth]{rotation_new}
\caption{Tracks of 3 PAPER-128 baseline classes shown in Fig. \ref{fig:AntPos}, here with the same respective colors. Tracks are traced out over half a sidereal day for frequency $\nu=0.15\text{\,GHz}$ and a phase center that passes through zenith. The top panel shows $uv$ tracks with crossings among each pair of baseline classes. Projection to the $vw$ plane in the bottom panel shows that tracks in $uvw$ space do not cross. }
\label{fig:Tracks}
\end{figure}

Equivalent baselines follow identical $uvw$ tracks. Traditionally, we can identify
redundancy of nearly equivalent baselines as crossings
of the $uv$ tracks, a 2 dimensional projection of Fig. \ref{fig:Tracks}. As we see in the top panel of Fig. \ref{fig:Tracks}, there would be many such crossings. However, there are several reasons that $uv$ track-crossing do not imply perfect redundancy. In fact, in our case $uv$ track-crossings are
not accurate enough for time offset determination, nor can it give estimate of the degree of redundancy.  The most obvious reason is that the 3 dimensional $uvw$ tracks do not actually cross, as is evident from the $vw$ projection in the bottom panel of Fig. \ref{fig:Tracks}. For drift-scan arrays, even a hypothetical crossing in $uvw$ space would not imply perfect redundancy. This shall become evident in the next section, after we develop a more general formalism that accounts for the point spread function of the finite beams, and estimate the degree of redundancy for general combination of baselines at a general time offset. Then, relation of track-crossing and redundancy for both drift-scan and tracking measurements will be explored in Appendix \ref{sec:appA}.

@@

%\caption{Example beam response of HERA (left) and PAPER (right) antennas. Both beams are Stokes $I$ polarization antenna voltage beams, at frequency of $\nu=150\text{\,MHz}$ and normalized to $1$ at zenith. The circles centered around zenith (center of beam) here are
%spaced 10 degrees apart. \label{fig:Beam}}
%\end{figure}

Power-spectrum measurements are typically taken from $\sim10$\,MHz centered around the corresponding redshift of interest (e.g. 150 \,MHz for z=9.5).

@@

Ultimately,
we would like to relate the data, collected with coordinates $\hat{\boldsymbol{s}}$
and $\nu$, to the power spectrum, written with 
$\boldsymbol{r}$ and $\boldsymbol{k}$, the cosmological position coordinate and wavenumber. To do so, we start by noticing that
\[
\begin{aligned}r & =\frac{c}{H_{0}}\int_{0}^{z}\frac{dz'}{E(z')},\\
 & \approx\frac{c}{H_{0}}\int_{0}^{z_{0}}\frac{dz'}{E(z')}-\frac{c(1+z)^{2}}{\nu_{21}H_{0}E(z)}\left(\nu-\nu_{0}\right),\\
 & \equiv X-Y\Delta\nu,
\end{aligned} \label{eq:r}
\]
where $\nu_{21}=1420$\,MHz is the 21\,cm transition rest frequency, $\nu_{0}$
a reference central frequency with corresponding redshift $z_{0}$,
and 
\[
E(z)=\sqrt{\Omega_{m}(1+z)^{3}+\Omega_{\Lambda}}.
\]
Inverting for $\nu$:
\begin{equation}
\nu=\frac{X-r}{Y}+\nu_{21}.\label{eq:nur}
\end{equation}

@@

Existing visibility based power-spectrum pipelines for redundant drift-scan  arrays relate the power-spectrum to the conjugate square of the visibilities \citep{delay-transform, paper32, Ali2015}. We would like to generalize such relations by relating the power-spectrum to the product of two visibilities from two arbitrary baselines and time offsets. 
The beam pattern of a baseline shifts relative to the sky as the Earth rotates. Here we choose to fix the sky, and denote the rotated coordinates
in the topocentric frame with the 3 dimensional rotation operator $\boldsymbol{\Gamma}$:

@@

With implicit bounds of integrals from $-\infty$ to $\infty$, we have:
\begin{widetext}
\begin{equation}
\begin{aligned} \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle
 & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),\\
 & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),\\
 & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),\\
 & \approx P(k_{\bar{b}, \tau})\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2} \delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),\\
 & = P(k_{\bar{b}, \tau})\int \frac{d^{3}r}{(X^2Y)^2}B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r})\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right],\\
 & = P(k_{\bar{b}, \tau})\int\frac{d\Omega d\nu}{X^{2}Y}B^{*}(\hat{\boldsymbol{s}},\nu)B(\boldsymbol{\Gamma} \hat{\boldsymbol{s}},\nu)\exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right],
\end{aligned}
\label{eq:main}
\end{equation}

@@

\begin{equation}
\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle\\
 & \qquad \approx P(k_{\bar{b}, \tau})\int\frac{d\Omega d\nu}{X^{2}Y}B^{*}(\hat{\boldsymbol{s}},\nu)B(\boldsymbol{\Gamma}\hat{\boldsymbol{s}},\nu)\\
 & \qquad \qquad \qquad \qquad \exp\left[i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right].
 \end{aligned}
\label{eq:final}
\end{equation}

@@

\begin{equation}
 \hat{P}(k_{\bar{b}, \tau}) \equiv \frac{V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)}{\Theta}, 
 \label{eq:opp}
\end{equation}
where the weight is defined as 
\begin{equation}
\Theta \equiv\int d\nu \Theta_{\nu}, 
\label{eq:Theta}
\end{equation}
with

@@

To first order, the physical origin of $\psi_\nu$ lies in the two visibilities having different phase centers. By default the correlators of a drift-scan array phase the two visibilities both to zenith at the same time. When they are cross-multiplied with a time lag, the visibilities must be rephased before delay transform to the account for the movement of the zenith. The effect of the phase thus roughly corresponds in a shift of the delay mode measured, and we should expect $\psi_{\nu}$ to be in first order a linear function of $\nu$:
\[
\psi_{\nu}\sim 2\pi\Delta\tau\nu.
\]
In Fig. \ref{fig:phi_nu} we show the peak phases for for a given baseline pair of PAPER128 (\{2,0:2,1\}), comparing the drift-scan phase dependence with that of the same baseline with hypothetical tracking-elements (Eq. \eqref{eq:Thetanu} and Eq. \eqref{eq:Thetanu_tracking}). In the drift-scan case, we indeed see a linear relation corresponding to a delay of $\Delta\tau\approx15.7$ns. In the tracking case, because of minimal zenith movement, only second order effects are observed. The origin of the second order effects can be seen as due to the $w$-term, or more precisely the fact that $uvw$ tracks do not cross when their 2 dimensional projections do. We refer the reader to Appendix \ref{sec:appA} for further explanation. In both cases, the full effects are encapsulated in the phase of $\Theta_\nu$ and can be thus determined empirically without any additional computation.

@@

\caption{Frequency dependent peak phases for PAPER-128 baseline pair \{2,0:2,1\}. Both the drift-scan case and a hypothetical tracking baseline are shown in solid lines. The first-order, linear effects are shown in dashed lines. We have also fixed the global phase in both cases to 0 at 0.15\,GHz. The drift-scan case exhibits linear behavior and conforms well to the first order effect of a shift in delay space due to the movement of zenith. }
\label{fig:phi_nu}
\end{figure}

To illustrate the effect of rephasing, we compare in Fig. \ref{fig:freqdiff} real parts of $\Theta_{\nu}$ of both equivalent and nearly equivalent baseline pairs for two channels: 0.145\,GHz and 0.155\,GHz. The top two panels have zero rephasing, and the bottom two are linearly rephased to a differential delay of $\Delta\tau\approx15.7$ns, as we determined above. The first and third panels show the equivalent baseline pairs \{2,0:2,0\}, and second and fourth panels show \{2,0:2,1\}. Summing over frequency without proper rephasing leads to destructive interference and signal loss. The wider the frequency profile, the more destructive the interference would be. Only after rephasing to the correct time offset for each baseline pairs can we constructively combine the frequency channels.

@@

So far for clarity and generality we have avoided the traditional formulation of radio astronomy in terms of the $uv$-plane. To gain some visual intuition of the formalism in Section \ref{sec:method}, we show explicit beam fringe patterns of the two baselines and their interactions. For this section, we relax into the flat sky approximation so that $(l,m,\sqrt{1-l^2-m^2})\equiv\hat{s}$ and examine a single frequency of $150$\,MHz.

At the time offset of maximum redundancy, we expect two beams to have to same fringe pattern in both frequency and phase. Due to the time delay, however, the beam centers would be slightly shifted with respect to each other. This we show in the top panels of Fig. \ref{fig:beamfringe}. The top left and middle panels show the beam-fringe
patterns (real parts) for baselines \{2,0\} and \{2,1\}, delayed by 0.0325 sidereal days.
The product of the beam fringe patterns in top right shows that the fringes
 cancel out as we expect. On the bottom of Fig. \ref{fig:beamfringe} we show the instantaneous $uv$ coverage of two baselines and their product. We see from the bottom right panel that the cross-multiplied baselines recover  power concentrated at $u\approx\frac{30\text{m}}{c/1.5\times10^8\text{Hz}}\approx15$. Note the bottom right panel is not the Fourier transform of the top right panel.

@@

\caption{PAPER128 beam-fringe patterns and their point-spread functions on the $uv$-plane for a frequency of $150$\,MHz. Selected baselines have 30 meter East-West separation, and 0 and 4 meter North-South separation, respectively. The top panels show the real parts of the beam-fringe patterns of the baselines (left and middle), and their conjugate product (right). The beam-fringe values are normalized such that the peak of the original beam is unity. 
The bottom left and middle panels show the peak-normalized $uv$ beam power of the two beam-fringe patterns. The product of the beam spreads displayed in the bottom right panel shows that power concentrated at $u\approx15$ is recovered. }
\label{fig:beamfringe}
\end{figure*}

@@

Bringing together the discussion from Section \ref{sec:method}, we present a numerical check of Eq. \eqref{eq:final}, including rephasing. To do so, we need to compare the amplitude and phase of the integral weight $\Theta$ 
for a pair of baselines with products of simulated visibilities of those baselines. We use 10 frequency channels evenly spaced from $145$\,MHz to $155$\,MHz for the comparison. We rephase one of the simulated visibilities, as well as computed $\Theta_{\nu}$ to the same delay as computed from the peak phase of the latter.

@@

Fig. \ref{fig:numerics} shows the real parts of the resulting degree of redundancy as a function of time offset, for equivalent baseline pairs (\{2,0:2,0\}) in the top panel and nearly equivalent (\{2,0:2,1\}) in the bottom. Shown in green for the two cases are the computed values of $\Theta$, both normalized to the peak of the equivalent case. The plot on the bottom is rephased with a delay of $\Delta\tau\approx15.7$ns as outlined in Section \ref{sec:rephs}. Since both y-axes are normalized to the same scale, we see the equivalent baseline reach maximal redundancy with 0 time offset, while the nearly equivalent pair reaches maximum redundancy of $\Theta\approx0.5$ at $dT\approx0.325$days.

Almost completely overlapping with the green curves are the cross-multiplied visibilities from a simple simulation, shown in blue. For the simulation, we populate random values of brightness temperature
on a healpix map \citep{Heal, HealPrimer} \footnote{We use functionalities in the python package AIPY for healpix mapping
as well as coordinate transforms. }. We rotate the baseline positions with
the appropriate rotation matrix, multiplying the sky by the primary beam to get the visibilities, for each
baseline\footnote{There are two obvious ways to achieve the rotation. One
can either fix the sky and rotate the baselines, or the other way
around. We found however, that we must not physically rotate the sky
map, for the numerical round-offs due to finite resolutions of the
map turns out to be significant. Thus we let the sky, represented
by the healpix map, be fixed, and rotate the baselines. }. For the nearly equivalent case, we rephase the delayed visibility (\{2,0\}) by the same delay $\Delta\tau\approx17$ns, followed by delay-transform. The resulting delay-space visibilities for the two baselines are then convolved
via the Fourier convolution theorem, to obtain values of the cross
correlation as a function of time offset.   We do this for both the equivalent (\{2,0:2,0\}) and nearly equivalent (\{2,0:2,1\}) case.
Since the blue and green curves overlap, we have verified that in this case Eq. \eqref{eq:final} is valid and the computed $\Theta$ can be used as weight to obtain the power spectrum.
